---
title: 'EDA and preprocessing'
author: 'Luke Adams'
date: '2021-06-24'
output: rmarkdown::github_document
---

**Setup**

```{r setup, include=TRUE, message=FALSE}
# setup
library(knitr)
knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE, warning = FALSE,
                      message = FALSE, echo = TRUE, dpi = 180, 
                      fig.width = 10, fig.height = 6)
library(tidyverse)
library(tidymodels)
theme_set(theme_minimal())
update_geom_defaults('rect', list(fill = 'midnightblue', alpha = 0.8))

library(DALEXtra)
library(lime)
library(GGally)     ## ggcor()

library(here)
library(skimr)
library(janitor)

library(caret)     # nearZeroVar
library(visdat)

# mosaic plot
library(matrixStats)
library(ggmosaic)
```

**Read data**

```{r, message=F}
# read data
data_00 <- read_csv(here('data', 'hmeq-data-raw.csv'))
```

## Basic EDA

-   5,960 individual loans

-   12 features with **REASON** and **JOB** being categorical

-   **BAD** is our response variable (1 = default, 0 = no default)

-   NAs are present

```{r}
# quick look at the data
glimpse(data_00, width = 90)
```

**Clean up column names**

```{r}
# rename columns for easier interpretability
# clean_names() converts columns to snake case by default
data_01 <-
  data_00 %>% 
  rename(default = BAD, loan_amt = LOAN, mort_due = MORTDUE, prop_val = VALUE, 
         loan_purpose = REASON, job = JOB, job_years = YOJ, num_derog = DEROG,
         num_delinq_credit_lines = DELINQ, credit_line_age_months = CLAGE, 
         num_credit_inq = NINQ, num_credit_lines = CLNO, debt_to_income = DEBTINC) %>% 
  clean_names()
```

**Convert data types**

-   **default**, **loan_purpose**, **job** to factor variables

-   Recode **default** factors (1 is yes, 0 is no)

```{r}
data_02 <- 
  data_01 %>% 
  modify_at(c('default', 'loan_purpose', 'job'), as.factor) %>% 
  mutate(default = recode(default, `1` = 'yes', `0` = 'no'))
```

Save model ready data

```{r}
glimpse(data_02, width = 90)
```

```{r}
# write_csv(data_02, here('data', 'hmeq-tidy.csv'))
```

**Detailed view**

-   1,189 / 5,960 loans defaulted (\~20%)

-   **loan_purpose** has 2 unique values, **job** has 6 unique values

-   Missing values are present in all variables except **default** and **loan_amt** with our lowest completion rate being **debt_to_income** at 78.7%

-   Numeric variables appear to be right skewed based off histograms

-   All numerics are greater than one and maximum values appear reasonable for each variable

```{r, }
skim(data_02) %>% 
  knit_print()
```

**Categoricals**

**loan_purpose** has 2 unique values - home improvement and debt consolidation

```{r}
data_02 %>% 
  select(loan_purpose) %>% 
  unique()
```

**job** has 6 uniques - office, sales, manager, executive, self-employed, other

```{r}
data_02 %>% 
  select(job) %>% 
  unique()
```

**Zero Variance Predictors**

No near zero variance predictors to be removed

```{r}
data_02 %>% 
  select_if(is.numeric) %>% 
  nearZeroVar(saveMetrics = TRUE) %>%
  rownames_to_column() 
```

## Visualizations

### Missing values

Logistic regression requires missing values to either be dropped or imputed

For categorical - we can either create a new category called "unknown" or can impute NAs - we'll come back to this

-   No clear or obvious pattern of missing values

-   Numeric missing values we will preprocess with a simple median or mean imputation as these are computationally efficient (may be worth it to compare finals results to models preprocessed with more advanced imputation methods)

```{r}
vis_dat(data_02)
```

### Response

-   Imbalanced response as we noticed earlier

```{r}
data_02 %>% 
  ggplot(aes(x = default)) +
  geom_bar()
```

```{r}
data_02 %>%
  group_by(default) %>%
  summarize(N = n(), pct = N / nrow(data_02))
```

### Features

#### Categorical

**loan_purpose, job**

```{r}
# loan_purpose category proportions
data_02 %>%
  tabyl(loan_purpose)
```

```{r}
# job category proportions
data_02 %>%
  tabyl(job)
```

#### Numeric

-   Numerics are right-skewed

```{r}
# distributions
data_02 %>%
  select(-default) %>%
  keep(is.numeric) %>%
  gather() %>%
  ggplot(aes(value)) +
    facet_wrap(~key, scales = 'free') +
    geom_density()
```

-   Outliers seem to be present (based on boxplot results)

```{r}
data_02 %>%
  select(-job, -loan_purpose, -default) %>%
  gather() %>%
  ggplot(aes(value)) +
    facet_wrap(~key, scales = 'free') +
    geom_boxplot()
```

#### Relationships

**loan_purpose, job x default**

Going to create a new category called unknown for each missing value - might be worth it to compare the final results to a model with imputed values (mode, KNN, bagged trees, etc.)

```{r}
data_02 %>% 
  select(default, loan_purpose, job) %>% 
  pivot_longer(loan_purpose:job) %>% 
  ggplot(aes(y = value, fill = default)) + 
  geom_bar(position = 'fill') +
  facet_wrap(vars(name), scales = 'free', ncol = 2) +
  labs(x = NULL, y = NULL, fill = NULL)
```

**loan_purpose x job**

```{r}
data_02 %>% 
  ggplot() + 
  geom_mosaic(aes(x = product(job, loan_purpose), fill = job))
```

**numeric x default**

```{r}
temp_2 <- data_02 %>%
  pivot_longer(-c(job, loan_purpose, default), names_to = 'variable')

ggplot(temp_2) +
 geom_boxplot(aes(x = factor(default), y = value)) +
 facet_wrap(~variable, scales = "free")
```

```{r}
data_02 %>% 
  select(default, loan_amt, mort_due, prop_val, num_derog, num_delinq_credit_lines, 
         debt_to_income, credit_line_age_months, num_credit_inq, num_credit_lines) %>% 
  ggpairs(columns = 2:10, aes(color = default, alpha = 0.5))
```

-   **mort_due** and **prop_val** have a correlation of 0.88, this might cause some instability in our logistic regression

-   this wouldn't be an issue if our objective was purely predictive but we are interested in interpretability as well

```{r}
data_02 %>% 
  select_if(is.numeric) %>% 
  ggcorr(label = TRUE, hjust = 0.89, label_round = 2)
```
