---
title: "Modeling"
author: "Luke Adams"
date: "6/24/2021"
output: rmarkdown::github_document
---

**Setup**

```{r setup, include=TRUE, message=FALSE}
# setup
library(knitr)
knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE, warning = FALSE,
                      message = FALSE, echo = TRUE, dpi = 180, 
                      fig.width = 10, fig.height = 6)
theme_set(theme_minimal())

library(tidyverse)
library(tidymodels)

library(themis)     # step_smote()
```

**Load data**

```{r}
# load data and convert to correct data types
hmeq_data <- 
  read_csv(here('data', 'hmeq-data-tidy.csv')) %>% 
  modify_at(c('default', 'loan_purpose', 'job'), as.factor)

glimpse(hmeq_data, width = 85)
```

# Split data

-   Stratified random sample split into a train, test set (3:1)

-   Resample the training set using repeated 10-fold cross-validation

```{r}
set.seed(9007)
hmeq_split <- initial_split(hmeq_data, strata = default)
hmeq_train <- training(hmeq_split)
hmeq_test <- testing(hmeq_split)

set.seed(9008)
hmeq_folds <- vfold_cv(hmeq_train, strata = default, repeats = 5)
```

# Recipe (preprocess)

Feature selection

-   drop correlated predictor

Subsampling

-   smote

Missing values

-   median for numeric

-   new category for nominal

Skewness

-   could transform variables but that would inhibit interpretability

Dummy variables

-   One hot encoding for nominal predictors

Feature engineering

-   interaction terms may increase performance but would increase complexity

```{r}
hmeq_rec <- 
  recipe(default ~ ., data = hmeq_train) %>% 
  step_meanimpute(all_numeric()) %>%
  step_unknown(all_nominal(), -default) %>%
  step_corr(all_predictors(), threshold = 0.85) %>% 
  step_dummy(loan_purpose, job) %>% 
  step_smote(default)
```

# Model specification

Going to train a basic logistic regression model for simplicity, efficiency and interpretability

```{r}
# model
(log_reg_spec <-
  logistic_reg() %>% 
  set_engine('glm'))
```

# Create workflow

Crucial step for keep preprocessing consistent when resampling model

```{r}
# workflow
(log_reg_wflow <- 
  workflow() %>% 
  add_recipe(hmeq_rec) %>% 
  add_model(log_reg_spec))
```
